Model based reinforcement learning at FERMI FEL

### 2020-11-16 main.tex
txs auto checkin

%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
% \usepackage{epic,eepic,pspicture,mathptmx,times,graphpap,siunitx, pgf, import}
% For algorithms
\usepackage{algorithm}
\usepackage{algpseudocode}
%\usepackage[ruled,vlined]{algorithm2e}

@@

\section{Deep Reinforcement learning}
Assume states $s$ the set of all states $S$ and actions $a$ all actions $a$ and rewards $r$ and the set of all rewards $R$ an initial state distribution $d_0$.
Goal of reinforcement learning is to find a $\pi(s)\mapsto a$, which is the solution of:
\begin{align}
\max J(\pi)  =  \mathbb E_{\tau \thicksim p_{\pi}}\left[\sum_{t=0}^{H}\gamma^t r(s_t,a_t)\right],
\end{align}
where $p_\pi = d_0(s_0)\prod_{t_0}^H\pi(a_t,s_t)\mathcal T(s_{t+1}|s_t,a_t)$ is the distribution of all trajectories $\tau := (s_0, a_0, s_1, a_1,...s_H,a_H)$ drawn by $\pi$ with a horizon $H$.
In the modern field of RL one distinguishes if the policy $\pi(s)\approx \pi_\phi(s)$ or the state-action value function $Q(s,a)$ is approximated using a high capacity function approximator, as e.g. a deep neural network. In the first case one speaks about policy gradient methods in the latter about approximate dynamic programming, which we now discuss.
\subsection{Approximate dynamic programming}
The state-value function $Q(s,a)$ is expressed as $Q_\phi(s,a)$, where $\phi$ denotes the parameters of the function approximator. By satisfying the Bellmann-optimality operator the optimal $Q^*(s,a)$ can be found:
\begin{align}
    \min_\theta \left(\vec Q_\theta - \mathcal{B}^*\vec Q_\theta\right)^2.\label{eq:minimize_bellmann_optimality}
\end{align}
This operator has a unique fixed point but is non-linear, due to the $\max$ - operator:
\begin{align}
 \mathcal{B}^*\vec Q_\theta(s_t,a_t) = r(s_t,a_t)+\gamma\max_a\left( Q_\theta(s_{t+1},a_t) - Q_\theta(s_t,a_t)\right).
\end{align}
The form of this equation can cause overestimation and other complications, when using a function approximator. Several methods exist which try to mitigate the problems \cite{}.
One way to reduce the effect is to take a simple analytical form of the $Q$-function.
\subsection{NAF}
If a specific quadratic form of the $Q$ function is assumed:
\begin{align}
     Q_\theta(s,a) = -\frac{1}{2}(a-\mu_\theta(s))P_\theta(a-\mu_\theta(s))^T+V_\theta(s).
\end{align}
 The optimal policy is given by:
\begin{align}
    \pi^*(s)=(\delta(a)-\arg\!\max_a Q_\theta(s,a)).
\end{align}
One modification, which is used in these tests is a twin network (weights denoted by $\tilde\theta$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
The maximum is given analytically as $\max_a Q(s,a) = V(s)$, hence from \cref{eq:minimize_bellmann_optimality}:
\begin{align}
	\min_\theta\left( (r(s_t,a_t)+\gamma \min_{\theta\prime,\tilde \theta} V_{\theta\prime}(s_{t+1}) - (1+\gamma) Q_\theta(s_t,a_t)\right)^2
\end{align}
$\theta\prime$ denotes the weights of a previous iteration.
%\subsection{New NAF network architecture}

\begin{algorithm}[ht]
\caption{On-policy policy gradient with Monte Carlo estimator \label{alg:pg}}
\begin{algorithmic}[1]
\State initialise $\theta_0$
%\For{iteration $k \in [0, \dots, K]$}
%\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
%\State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
%\State fit $b(\bs_t)$ to $\{\return{i,t}\}$ \Comment{use constant $b_t = \frac{1}{N}\sum_i \return{i,t}$, or fit $b(\bs_t)$ to $\{\return{i,t}\}$}
%\State compute $\hat{A}(\bs_{i,t},\ba_{i,t}) = \return_{i,t} - b(\bs_t)$
%\State estimate $\nabla_{\theta_k} J(\policy_{\theta_k}) \approx \sum_{i,t} \nabla_{\theta_k} \log \policy_{\theta_k}(\ba_{i,t} | \bs_{i,t}) \hat{A}(\bs_{i,t},\ba_{i,t})$
%\State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
%\EndFor
%\end{algorithmic}
\end{algorithm}

>>>>

%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
% \usepackage{epic,eepic,pspicture,mathptmx,times,graphpap,siunitx, pgf, import}
% For algorithms
%\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[ruled,vlined]{algorithm2e}

@@

\section{Deep Reinforcement learning}
Assume states $s$ the set of all states $\states$ and actions $a$ all actions $\actions$ and rewards $r$ and the set of all rewards $\rewards$ an initial state distribution $d_0$.
Goal of reinforcement learning is to find a $\pi(s)\mapsto a$, which is the solution of:
\begin{align}
\max J(\pi)  =  \mathbb E_{\tau \thicksim p_{\pi}}\left[\sum_{t=0}^{H}\gamma^t r(s_t,a_t)\right],
\end{align}
where $p_\pi = d_0(s_0)\prod_{t_0}^H\pi(a_t,s_t)\mathcal T(s_{t+1}|s_t,a_t)$ is the distribution of all trajectories $\tau := (s_0, a_0, s_1, a_1,...s_H,a_H)$ drawn by $\pi$ with a horizon $H$.
In the modern field of RL one distinguishes if the policy $\pi(s)\approx \pi_\phi(s)$ or the state-action value function $Q(s,a)$ is approximated using a high capacity function approximator, as e.g. a deep neural network. In the first case one speaks about policy gradient methods in the latter about approximate dynamic programming, which we now discuss.
\subsection{Approximate dynamic programming}
The state-value function $Q(s,a)$ is expressed as $Q_\phi(s,a)$, where $\phi$ denotes the parameters of the function approximator. By satisfying the Bellmann-optimality operator the optimal $Q^*(s,a)$ can be found:
\begin{align}
    \min_\theta \left(\vec Q_\theta - \mathcal{B}^*\vec Q_\theta\right)^2.\label{eq:minimize_bellmann_optimality}
\end{align}
This operator has a unique fixed point but is non-linear, due to the $\max$ - operator:
\begin{align}
 \mathcal{B}^*\vec Q_\theta(s_t,a_t) = r(s_t,a_t)+\gamma\max_a\left( Q_\theta(s_{t+1},a_t) - Q_\theta(s_t,a_t)\right).
\end{align}
The form of this equation can cause overestimation and other complications, when using a function approximator. Several methods exist which try to mitigate the problems \cite{}.
One way to reduce the effect is to take a simple analytical form of the $Q$-function.
\subsection{NAF}
If a specific quadratic form of the $Q$ function is assumed:
\begin{align}
     Q_\theta(s,a) = -\frac{1}{2}(a-\mu_\theta(s))P_\theta(a-\mu_\theta(s))^T+V_\theta(s).
\end{align}
 The optimal policy is given by:
\begin{align}
    \pi^*(s)=(\delta(a)-\arg\!\max_a Q_\theta(s,a)).
\end{align}
One modification, which is used in these tests is a twin network (weights denoted by $\tilde\theta$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
The maximum is given analytically as $\max_a Q(s,a) = V(s)$, hence from \cref{eq:minimize_bellmann_optimality}:
\begin{align}
	\min_\theta\left( (r(s_t,a_t)+\gamma \min_{\theta\prime,\tilde \theta} V_{\theta\prime}(s_{t+1}) - (1+\gamma) Q_\theta(s_t,a_t)\right)^2
\end{align}
$\theta\prime$ denotes the weights of a previous iteration.
%\subsection{New NAF network architecture}

\begin{algorithm}[ht]
\caption{On-policy policy gradient with Monte Carlo estimator \label{alg:pg}}
\begin{algorithmic}[1]
\State initialise $\theta_0$
\For{iteration $k \in [0, \dots, K]$}
%\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
%\State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
%\State fit $b(\bs_t)$ to $\{\return{i,t}\}$ \Comment{use constant $b_t = \frac{1}{N}\sum_i \return{i,t}$, or fit $b(\bs_t)$ to $\{\return{i,t}\}$}
%\State compute $\hat{A}(\bs_{i,t},\ba_{i,t}) = \return_{i,t} - b(\bs_t)$
%\State estimate $\nabla_{\theta_k} J(\policy_{\theta_k}) \approx \sum_{i,t} \nabla_{\theta_k} \log \policy_{\theta_k}(\ba_{i,t} | \bs_{i,t}) \hat{A}(\bs_{i,t},\ba_{i,t})$
%\State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
\EndFor
\end{algorithmic}
\end{algorithm}
